Explorations to Optimize the Parareal Algorithm \\to Solve ODEs in Given Applications

### 2014-12-09 paper.tex

Another obstacle to trying to paralleliize numerical methods is that the methods
are inherently serial in time---in that evaluation of the next time $n+1$
depends on the previous values, say those at time $n-1$ and $n$. This setup
does not allow for parallelism.

 An alternative paradigm is a parallelism by space, where the result vector is
 divided up to different processors.  In this way, the

In general, the

Due to the more theoretical nature of the parareal method, we look to describe
its stability and error numerically.

@@

Show that this parareal method will approach, with large enough $k$ (enough
iterations) to approach the error of the fine method. However, with too large
of a k and a lower quality factor (ratio of fine to coarse), the time for the
parareal could potentially take longer than the direct serial computation.

\subsection{Visualization Diagram}
bksim TODO

@@

We looked

We see that the speedups are

Note that increasing the iterations

@@

TODO team?

\subsection{Comparison to a Forward Euler Parallelized by Space}

>>>>

Another obstacle to trying to paralleliize numerical methods is that the methods are inherently serial in time - in that evaluation of the next time $n+1$ depends on the previous values, say those at time $n-1$ and $n$.  This setup does not allow for parallelism.  The only way to incorporate parallelism is to create schemes that have parallelizable components - such as an update or a refinement to a serial portion.  As will be discussed and analyzed in the metodology section, one such scheme would be the parareal algorithm.  For an overview, the parareal algorithm first does a coarse (low-order and fast) approximation, then seeks to iteratively correct using smaller, more refined numerical methods done in parallel from approximated starting points approximated by the coarse solution.

An alternative paradigm to applying parallelism to numerical methods is the parallelism by space paradigm, where the result vector space is divided up to different regions and sent to different responsible processors.  In this way, each processor sees a similar but smaller-scale problem - weak scaling.  This method is expected to outperform the parallel in time paradigm simply because the task is somewhat embarassingly parallel, being able to be naturally divided into smaller but identical in method problems.  There is no serial component in this algorithm and should be exhibit both strong and weak scaling.

In general, the division in space paradigm is more intuitive to parallelize and will often lead to greater speedups because there is no serial component.  Yet it is worth noting that knowing how to divide the result vector requires knowledge of problem-specific details.  Thus, this approach lacks the generalisability and theoretical derivaions that schemes that parallelize in time can offer.

\subsection{Visualization of Scheme}

@@

Show that this parareal method will approach, with large enough k (enough iterations) to approach the error of the fine method.  However, with too large of a k and a lower quality factor (ratio of fine to coarse), the time for the parareal could potentially take longer than the direct serial computation.

@@

We looked at both $e^x$ as well as TODO WHAT OTHER SCHEME using GET PARAMETERES

Varying Iterations:

Varying Quality Factor:
We define the quality factor in this case for how many times smaller the Euler method used for $g_{fine}$ is compared to the Euler method used for $g_{\Delta t}$

Changing the ODE

@@

TODO team

IDEALLY WE SHOW THE SAME GRAPHS AS THE ABOVE WITH ALL THE VARIATIONS TESTED

\subsection{Parallelism by Space Paradigm}
